---
title: "01-basic-ggplot-plotly"
author: "Fleur.Lecarpentier"
date: "11 novembre 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This markdown show how to confugure ggplot, ggplotly and plotly for a basic chart as bar chart.


```{r, message = FALSE}
library(ggplot2)
library(plotly)
library(dplyr)
```


GOAL : Create a toy dataset

```{r}
## data
  set.seed(999)
  DF002 <- data.frame( type = sample(c("blue", "red"), replace =T, size = 200),
                       x = sample((1:50), replace =T, size = 200))
  DF002$y <- log(DF002$x) + rnorm(1)

```


GOAL : Show how to deal with geom_ object : 
  default geom_bar : 
  - aes : 1 x-axis 
  - stat : count => count the # of observation per x
  - position : stack
  
```{r}
## plot - implicit default
   ggplot(DF002) + 
   geom_bar(aes(x = x, fill = type), stat = "count", position = "stack")

## same
   ggplot(DF002) + 
   geom_bar(aes(x = x, fill = type))
```


GOAL : Show how to deal with geom_object : changing aesthetic, stat and position
  user-defined geom_bar : 
  - aes : 2 axis 
  - stat : identity => plot the values inside y
  - position : dodge

```{r}   
## plot - change stat and position : 
  ggplot(DF002) + 
  geom_bar(aes(x = x, y = y, fill = type), stat = "identity", position = "dodge")
```


GOAL : Show how to deal with geom : add several geom_layer : ONLY IF HOMOGENEOUS Y-AXIS !

```{r}   
## plot - change stat and position : 
  ggplot(DF002) + 
  geom_bar(aes(x = x, y = y, fill = type), stat = "identity", position = "dodge") + 
  geom_line(aes(x = x, y = y), color = "black", size = 1, linetype = "dotted")

```


GOAL : create a numerical fake dataset for demo

```{r}
DF003 <- data.frame(color = sample((1:2), size = 20, replace = T),
                   price = sample((1:2), size = 20, replace = T))
```




GOAL : understand stat with ggplot, included internar computed variables for each stat : 
stat_count creates 2 internal variables : ..count.. and ..prop.. : 
  
The 2 following charts are exactly the same : only label countaining computed stat changes
- Nota : aesthetic shared by all layer can be factorized in ggplot(...)
- note by the way the use of facet_grid(facets = ~[var]) to plot chart side by side

```{r}
## Using ..count..
  ggplot(DF003, aes(x = color, fill=factor(price))) +     
    geom_bar() +    
    geom_text(aes(label = ..count..),   #use  prop.table(..count..) to get global %
              stat= "count",
	            position = position_stack(vjust=0.5)) + 
  facet_grid(~factor(price))

##using ..prop.. (BEWARE : ..prop.. calculate proportion per group !)
  ggplot(DF003, aes(x = color, fill=factor(price))) +     
    geom_bar() +    
    geom_text(aes(label = scales::percent(..prop..)),   # library(scales) provide nice formats
              stat= "count",
	            position = position_stack(vjust=0.5)) + 
  facet_grid(~factor(price))

```

With plotly


```{r}
 DF002 %>%
    group_by(x, type) %>%
    summarise(nb = n()) %>%
  plot_ly() %>% 
    add_trace(x = ~x, y = ~nb, color = ~type, type = "bar", 
            colors = c("red" = "lightcoral", "blue" = "steelblue"),
            hoverinfo = 'text', 
            text = ~paste('Type : ', type,'<br> x : ', x, '// y :', nb)) %>%
    layout(barmode = 'stack')

```



