---
title: "02-NewBusiness-NetInflow"
author: "Fleur.Lecarpentier"
date: "11 novembre 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) ## add a comment
```


```{r, message = FALSE}
library(ggplot2)
library(ggpubr)
library(dplyr)
library(scales)
library(plotly)
library(data.table)
```


## Cancellation and NB rate using geom_line : 

1. Create Dataset
```{r}

## data
  reportDF <- read.csv2(file = "./datasets/report2017.csv")
  reportDF$date <- as.Date(reportDF$date)
  reportDF$portfolio <- as.numeric(reportDF$portfolio)


  DFagg <- as.data.table(
          reportDF %>%
           group_by(LOB, network, date) %>%
           summarize(NB = sum(newBusiness),
                     canc = sum(cancellation),
                     portfolio = sum(portfolio)) %>%
           mutate(NB.rate = NB/portfolio,
                  canc.rate = canc/portfolio,
                  netInflow = as.numeric(NB - canc)) %>%
           select(-c(NB, canc))
  )

  DF.tidy <- DFagg %>% 
             tidyr::gather(KPI, value, -date, -LOB, -network) %>% 
             filter(network == "Agent") %>% 
             as.data.table()

```


2. Call ggplot on the tidy dataframe
```{r}
## generic theme  :
  myTheme <-  function(x) { theme(axis.text.x = element_text(angle = 90),                  
                                  axis.title.x = element_blank(),
                                  axis.title.y = element_blank(),
                                  legend.position = "right")
  }


 a1 <-   ggplot(DF.tidy[KPI %in% c("NB.rate", "canc.rate")]) +                        # data
         geom_line(aes(x=date, y=value, color = KPI)) +                               # geoms
         facet_wrap(~LOB, dir = "h") +                                                # facet for LOB
         scale_y_continuous(labels = scales::percent ) +                              # Scale Y
         scale_x_date(date_labels = "%b %y", date_breaks = "3 months") +              # Scale X
         scale_colour_manual(name="KPI's", 
                             values=c("NB.rate" = "green", "canc.rate" = "red"),
                             label = c("NB.rate"="New Business", "canc.rate"="Cancellation")) +
        myTheme() + ggtitle("Evolution of New Business/Cancellation rate per Line of Business")   # Theme
 a1
 
```
 
 
### SOLUTION 2 : Less adviced : using dataframe as it is

1. same data
2. ggplot on the untidy dataframe
```{r}
 b1 <-  ggplot(DFagg[network == "Agent"]) +                                           # data
        geom_line(aes(x=date, y=NB.rate, color = "NB")) +                             # geom NB
        geom_line(aes(x=date, y=canc.rate, color = "Canc")) +                         # geom Cancellation
        facet_wrap(~LOB, dir="h") +                                                   # facet for LOB
        scale_y_continuous(labels = scales::percent ) +                               # Scale Y
        scale_x_date(date_labels = "%b %y", date_breaks = "3 months") +               # Scale X
        scale_color_manual(name="KPI's", 
                            values= c(NB = "#00FF00", Canc = "#FF0000"),
                            label = c(NB = "New Business", Canc = "Cancellation")) +
        myTheme() + ggtitle("Evolution of New Business/Cancellation rate per Line of Business")   # Theme
b1    

```


### SOLUTION 3 : Solution 2 wrapped in a function 

Note that to map ggplot in a function, some adjustments need to be done : 
- aes => aes_string if the function variables appear in aes()
- Quoted string need to be wrapped with shQuote()
- as.formula for facet options

```{r}

   plot.NB.canc <- function(DF, x.date, y.nb, y.canc, grp) {
        ggplot(DF) + 
          geom_line(aes_string(x=x.date, y=y.nb, color = shQuote("NB"))) +
          geom_line(aes_string(x=x.date, y=y.canc, color = shQuote("Canc"))) +
          facet_wrap(as.formula(paste0("~",grp)), dir="h") +
          scale_y_continuous(labels = scales::percent ) +
          scale_x_date(date_labels = "%b %y", date_breaks = "3 months") +
          scale_color_manual(name="KPI's", 
                              values= c(NB = "#00FF00", Canc = "#FF0000"),
                              label = c(NB = "New Business", Canc = "Cancellation")) +
          myTheme() + ggtitle("Evolution of New Business/Cancellation rate per Line of Business")   
      }    
   c1 <- plot.NB.canc(DFagg[network == "Agent"], "date", "NB.rate", "canc.rate", "LOB")
   c1
```


## Adding NetInflow bar chart and combine all.

=> Here follows the global solution (I choose the tidy method embedded in a function)


```{r}
## generic theme  :
  myTheme <-  function(x) { theme(axis.text.x = element_text(angle = 90),                  
                                  axis.title.x = element_blank(),
                                  axis.title.y = element_blank(),
                                  legend.position = "right")
  }

## tidy dataframe
  DF.tidy <- DFagg %>% tidyr::gather(KPI, value, -date, -LOB, -network) %>% filter(network == "Agent") %>% as.data.table()


## function 1 : line 
 g1.f <- function(df, grp)  {ggplot(df) +                                                           # data
                       geom_line(aes(x=date, y=value, color = KPI)) +                               # geoms
                       facet_wrap(as.formula(paste0("~",grp)), dir="h") +                           # facet for LOB
                       scale_y_continuous(labels = scales::percent ) +                              # Scale Y
                       scale_x_date(date_labels = "%b %y", date_breaks = "3 months") +              # Scale X
                       scale_colour_manual(name="KPI",                                              # Scale Color
                                           values=c("NB.rate" = "green", "canc.rate" = "red"),
                                           label = c("NB.rate"="New Business", "canc.rate"="Cancellation")) +
                      myTheme()                                                                     # Theme
 }
 ## function 2 : bar  
 g2.f <- function(df, grp)  {ggplot(df) +                                                           # data
                       geom_bar(aes(x=date, y=value, fill = KPI),                                   # geoms
                                stat = "identity", position = "dodge") +                            # 
                       facet_wrap(as.formula(paste0("~",grp)), dir="h") +                           # facet for LOB
                       scale_y_continuous(labels = scales::comma ) +                                # Scale Y
                       scale_x_date(date_labels = "%b %y", date_breaks = "3 months") +              # Scale X
                       scale_fill_manual(name = "KPI",
                                         values=c("netInflow" = "khaki"),                           # Scale Fill
                                         label = c("netInflow"="Net Inflow")) +
                      myTheme()                                                                     # Theme
 }
 
 
 g1 <- g1.f(DF.tidy[KPI %in% c("NB.rate", "canc.rate")], "LOB")
 g2 <- g2.f(DF.tidy[KPI == "netInflow"], "LOB") 

ggpubr::ggarrange(g1, g2, align = "v", nrow = 2)


```

With ggplotly : 


```{r}
  subplot(ggplotly(g1, tooltip = c("date", "value")) , 
          ggplotly(g2, tooltip = c("date", "value")) %>%
                   layout(xaxis = list(side ="top", tickangle = -90), 
                          xaxis2 = list(side ="top",tickangle = -90)), 
          nrows =2, shareX = T, margin=.08)
  
  #plotly_json(p1)  # to understand how ggplotly object are built from ggplot

```

And With plotly


Technically feasible but not very lisible : it is adviced to create several plots.


```{r}
# choose your colors per LOB for instance
  cols <- data.table(v = levels(DFagg$LOB), colors = c("steelblue", "#dee5e4"))
  
  plotly.nbcc <- function(df, ntw, lob, legend) {
    
        # NB and Cancellation
        p.nbcc <- plot_ly(df[network == ntw & LOB == lob]) %>%
            add_lines(x = ~date, y = ~canc.rate*100, 
                      name = '% Cancellation',  
                      line = list(color = "red", dash = "dot"),
                      showlegend = legend,
                      hoverinfo = 'text', 
                      text = ~paste('% Cancellation : ', round(canc.rate*100, 1), "% ", '<br> (', ntw, '-', lob, ')')) %>%
            add_lines(x = ~date, y = ~NB.rate*100, 
                      name = '% NewBusiness',  
                      line = list(color = "green"),
                      showlegend = legend,
                      hoverinfo = 'text', 
                      text = ~paste('% New Business : ', round(NB.rate*100, 1), "% ", '<br> (', ntw, '-', lob, ')')) %>%
            layout(xaxis = list(title = ""),
                   yaxis = list(ticksuffix = '%', 
                                tickfont = list(size = 8)))
  }
  
  plotly.ni <- function(df, ntw, lob, legend) {
      # Net Inflow
      p.ni <- plot_ly(df[network == ntw & LOB == lob],
                x = ~date, y = ~netInflow, 
                type = 'bar', 
                name = "Net Inflow",
                marker = list(color = cols[v==lob, colors] ) ,
                showlegend = legend,
                opacity = .6,
                hoverinfo = 'text', 
                text = ~paste('Net Inflow : ', comma(netInflow), '<br> (', ntw, '-', lob, ')')) %>%
          layout(xaxis = list(tickangle = -90,
                              side = "top",
                              title = "",
                              autotick = F, 
                              dtick = "M2",
                              tickfont = list(size = 8)),
                 yaxis = list(tickfont = list(size = 8)))

 }
  
  #eval(plotly.nbcc(DFagg, "Agent", "Motor", T))
  #eval(plotly.ni(DFagg, "Agent", "Motor", T))
  
  
```

And see the results 

```{r}


 ## paste plots together
 plotly.KPI <- function(df, ntw, lob, legend) {
 
    subplot(plotly.nbcc(df, ntw, lob, legend), 
            plotly.ni(df, ntw, lob, legend), 
            nrows = 2, shareX = TRUE, margin = .08) %>%
    layout( title = paste("Net Inflow, New Business & Cancellation rates  : <b>", ntw, "-", lob, "</b>"),
            margin = list(b=70),
            legend = list( x = 1.05,                        # avoid overlap of legend and y-axis
                           y = 0.5, yanchor = "center") )   # valign
  
 }

 plotly.KPI(DFagg, "Agent", "Motor", T)
 
```


Use subplot to gather charts :

```{r}

   subplot(plotly.nbcc(DFagg, "Agent", "Motor", T), 
           plotly.nbcc(DFagg, "Direct", "Motor", F),
           plotly.ni(DFagg, "Agent", "Motor", T), 
           plotly.ni(DFagg, "Direct", "Motor", F), 
            nrows = 2, shareX = TRUE,shareY = T, margin = .08) %>%
    layout( title = paste("Net Inflow, New Business & Cancellation rates  : <b> Agent vs Direct - Motor</b>"),margin = list(t = 40),
            legend = list(x = 1.05,                        # avoid overlap of legend and y-axis
                           y = 0.5, yanchor = "center") )   # valign
  
```
