---
title: "04-Choropletha-maps"
author: "Fleur.Lecarpentier"
date: "11 novembre 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message = FALSE}
library(ggplot2) # install from hadley/ggplot2
library(dplyr)
library(RColorBrewer)
library(rgdal)
library(rgeos)
library(maptools)
library(leaflet)
library(ggmap)  # install.packages("ggmap", type = "source")
library(broom)
```

Get Data

```{r}
 France <- readOGR(dsn=paste0(getwd(), "/datasets/maps"), layer="FRA_dept")
 
```


#### with basic ggplot 

Let's use our polygons map to show statistics
By default, polygons id refer to @data rows. 

```{r}
## add an idea in France@data to merge with datasets
  France@data$id <- as.character(rownames(France@data))


## broom::tidy helps to create a dataframe using all polygons informations. 
## It lost @data, which need to be re-merged

  poly.data <- broom::tidy(France)
  
  poly <- left_join(poly.data, France@data, by = "id")

  
## plot a map with an automatic gradient fill
  ggplot(poly) + geom_polygon(aes(x=long, y=lat, group = group, fill = CR), color = "black", size=.2) + 
    scale_fill_gradientn(colors = RColorBrewer::brewer.pal(5,"RdYlBu")[5:1]) +
    coord_map()

#display.brewer.all(5)

```


#### adding Google map background ( !!need Web access)

```{r}
## centroid coordinates of each regions
  centr <- gCentroid(France)@coords
  centr.Paris <- gCentroid(France[France@data$DEPT_ID==75,], byid = TRUE)@coords

## get map (require internet access)  
  frMap <- get_map(location = centr, zoom = 6)


## then let’s plot our polygons. As both objects share the same CRS, layers fit exactly
  g <- ggmap(frMap) +
    geom_polygon(data = poly, 
                 aes(x=long, y=lat, group = group, fill = CR), color = "black", size=.2, alpha = .5) +      
    coord_map() + 
    scale_fill_gradientn(colors = RColorBrewer::brewer.pal(5,"RdYlBu")[5:1])
  g  

```



### example

Let's add information to our data-slot and plot it with a discrete scale.

1. Create a fake dataframe of 96 rows (= nb of polygons) containing 
- polygons id 
- a variable named "inhts" taking randomly values in "0-10k", "10-50k", "50-100k", "100k +"
(tips : you can use ```sample```)

2. Plot it on France map

```{r}
## random df
 inht.df <- data.frame(id = unique(poly$id), 
                      inhts = sample(c("0-10k", "10-50k", "50-100k", "100k +"), 
                                     size = length(unique(poly$id)), replace = T))

## merge
  poly2 <- left_join(poly, inht.df, by = "id")
  
  # in addition, I add it to the data-slot of our map
  France@data <- left_join(France@data, inht.df, by = "id")

## plot
  
 # Choose your palette
  display.brewer.all()
  
 # plot map
  ggmap(frMap) +
  geom_polygon(data = poly2, 
               aes(x=long, y=lat, group = group, fill = inhts), color = "black", size=.2, alpha = .5) +    
  coord_map() +
  scale_fill_brewer(palette = "YlOrRd")


```



#### with leaflet


```{r}

# color palette : 5 levels from Red to Green

  bins <- quantile(France@data$CR, c(0, .25, .5, .75, 1))
  pal <- colorBin("RdYlBu", domain = France@data$CR, bins = bins,
                           na.color = "grey40", reverse = T)


# finally leaflet map
 l <-  leaflet(options = leafletOptions(minZoom = 5, maxZoom = 8)) %>% 
        addTiles() %>% 
        setView(centr[1], centr[2], zoom = 5) %>% 
        addLegend(pal = pal, values = round(France@data$CR, 1), 
                  opacity = 0.7, position = "bottomright", title = "Combined Ratio")
 
 
 l %>%  addPolygons(data=France, weight = 1, 
                    fill = ~CR, fillColor = ~pal(CR),
                    opacity=1, fillOpacity = 0.6, color=grey(0.5),
                    ## USE POPUP
                    popup = ~as.character(
                        paste(DEPT_ID, DEPT, "<br>", "Combined Ratio =", round(CR, 2)))
                    ) 

 
```



You can easily adjusy hover options : in the following example, Popup is replaced by higlights

```{r}
 

 l %>% addPolygons(data=France, weight = 1, 
                    fill = ~CR, fillColor = ~pal(CR),
                    opacity=1, fillOpacity = 0.6, color=grey(0.5),
                    ## HIGHLIGHTS instead of POPUP
                    highlight = highlightOptions(
                         weight = 1,
                         color = "white",
                         dashArray = "",
                         fillOpacity = 0.7,
                         bringToFront = TRUE),
                    label = ~as.character(
                      paste(DEPT_ID, DEPT, "Combined Ratio =", round(CR, 2)))
              )  


```