---
title: "05-GPS-maps"
author: "Fleur.Lecarpentier"
date: "11 novembre 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message = FALSE}
library(ggplot2) # install from hadley/ggplot2
library(RColorBrewer)
library(dplyr)
library(rgdal)
library(rgeos)
library(leaflet)
library(ggmap)  # install.packages("ggmap", type = "source")

```

```{r}
  France <- readOGR(dsn=paste0(getwd(), "/datasets/maps"), layer="FRA_dept")

## centroid coordinates of each regions
  centr <- gCentroid(France)@coords
  centr.Paris <- gCentroid(France[France@data$DEPT_ID==75,], byid = TRUE)@coords

## get map (require internet access)  
  frMap <- get_map(location = centr, zoom = 6)


```

#### Ploting dots (set of GPS) _______________________________________________________________________


```{r}

  compet <- read.csv2(file = "./datasets/competitors.csv", stringsAsFactors = F)
  compet$competitor <- factor(compet$competitor)
  compet$lat <- as.numeric(compet$lat)
  compet$long <- as.numeric(compet$long)

  str(compet)
  
```



#### Using ggplot

If we take exactly the same previous example with polygon centroid instead of polygon itself : 
=> use geom_point

In addition we can test it with more data points and plot a density : 

```{r}

## points
  ggmap(frMap) +
  geom_point(data = France@data, aes(x=long_ctd, y=lat_ctd, color = CR)) +
  scale_color_gradientn(colors = RColorBrewer::brewer.pal(5,"RdYlBu")[5:1])


## density
  ggmap(frMap) +
  geom_point(data = compet, aes(x=long, y=lat), position = "jitter", alpha = .5)
  
  
  ggmap(frMap) +
  stat_density_2d(data = compet, aes(x=long, y=lat, fill = ..level..,  alpha =..level..), 
                  size = 1, bins = 16, geom = "polygon") +
  scale_fill_gradient(low = "green", high = "red") 


```



#### Using Leaflet

Plotting GPS coordinates => rapidly feel the need of zoom/dezoom, and globally dynamic feature.
A great javascript-based library for maps is leaflet : 


1. Create icon list
```{r}

## your own icons
  competitor.ico <- iconList(
      AXA = makeIcon(iconUrl = "img/axa.png", iconWidth = 20, iconHeight = 20),
      Groupama = makeIcon(iconUrl = "img/groupama.png", iconWidth = 20, iconHeight = 20),
      Allianz = makeIcon(iconUrl = "img/allianz.png", iconWidth = 20, iconHeight = 20)
)

```


2a. Plot GPS on a map (using our own icons :) )
```{r}

## simple dots
  leaflet(compet, options = leafletOptions(minZoom = 5, maxZoom = 18)) %>% 
    addTiles() %>%
    setView(centr.Paris[1], centr.Paris[2], zoom = 12) %>% 
    addMarkers(lng = ~long, lat = ~lat, icon = ~competitor.ico[competitor], 
               popup = ~as.character(paste(activity_label, "<br>", adr.street, "<br>", adr.city)))
    
```


2b. We can use js libraries like markercluster
```{r}
## markercluster
  leaflet(compet, options = leafletOptions(minZoom = 5, maxZoom = 18)) %>% 
    addTiles() %>%
    setView(centr[1], centr[2], zoom = 5) %>% 
    addMarkers(clusterOptions = markerClusterOptions() )
```




