---
title: "15-animate-maps"
author: "Fleur.Lecarpentier"
date: "11 novembre 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message = FALSE}
library(ggplot2)
library(plotly)
library(gganimate)
library(tidyr)
library(dplyr)
library(rgdal)
library(lubridate)
library(data.table)
library(scales)
```



#### Automated plots : 

This paragraph show how to animate plots (to illustrate evolution for instance)
In a first step, we will plot Calls per type on a Seattle map :
- using dots
- using density

then we'll show how to animate these plots over time with the special aesthetic ```frame```


Get data

```{r}
  load(file = "./datasets/Call911.Rdata", verbose = T)
  Seattle <- readOGR(dsn=paste0(getwd(), "/datasets/maps"), layer="Neighborhoods")

```


data prep :  First fortify all coordinates into a dataframe to use with ggplot

```{r}
  Seattle.points <- fortify(Seattle, region = "OBJECTID")
  
  ggplot(Seattle.points, aes(long, lat, group = group)) +
    geom_path() +
    coord_map()
```



data prep: Then adjust our Call 911 data frame to limit the number of Types and add year

```{r}
  ## See types and aggregate less represented
  list_of_types <- Call911.short %>% group_by(Type) %>% 
    summarise(nb = n()) %>% 
    arrange(desc(nb)) %>% 
    mutate(perc = scales::percent(nb/sum(nb))) %>%
    filter(nb/sum(nb)<.02) %>% 
    select(Type) %>% 
    unique() %>% unlist()


    ## change types 
    Call911.short2 <- Call911.short %>% 
                      mutate(Type = ifelse(Type %in% list_of_types, "Others", Type)) %>%
                      mutate(year = year(date), 
                             month = paste(month(date), months(date, abbreviate=T), sep="-")) %>% 
                      as.data.table()

    
    ## test
    Type.perc <- function(Y) {
          (call.lab <- Call911.short2 %>% filter(year %in% Y) %>% group_by(Type) %>% 
          summarise(nb = n()) %>% 
          arrange(desc(nb)) %>% 
          mutate(perc = nb/sum(nb))
        )
    }
    Type.perc(2017)
 
       
    ## add factor 
    Call911.short2$Type <- factor(Call911.short2$Type, levels = unlist(Type.perc((2010:2017))$Type))

```

plot : finally plot the result

```{r}

  Call.g1 <- function(Y) {ggplot(Call911.short2[year %in% Y]) +
              geom_point( aes(x=Longitude, y=Latitude, frame = year, color = Type), size = 0.01) +
              geom_path(data=Seattle.points, aes(long, lat, group = group), colour = "#d3d3d3") +
              ggthemes::theme_map() +
              theme(plot.background = element_rect(fill = "black"),
                    legend.background = element_rect(fill = "black"),
                    legend.position = "right",
                    legend.key = element_rect(fill = "black"),
                    title = element_text(color = "white", size =12),
                    legend.text = element_text(color = "white")) +
              guides(colour = guide_legend(override.aes = list(size=15*Type.perc(Y)$perc ))) +
              coord_map() +
              xlim(Seattle@bbox[1,1], Seattle@bbox[1,2]) +
              ylim(Seattle@bbox[2,1], Seattle@bbox[2,2]) +
              scale_color_viridis_d(option = "plasma", labels = paste(Type.perc(Y)$Type, percent(Type.perc(Y)$perc))) 
  }

# 1 chart
  Call.g1(2017)

```




And now let's animate : 

```{r, fig.show="animate"}

  gganimate(Call.g1(2010:2017)) 

```

Faceting : 

```{r}

  Call.g1(2017) + facet_wrap(facet = ~Type)

```

```{r, fig.show='animate'}

  gganimate(Call.g1((2010:2017)) + facet_wrap(facet = ~Type)) 

```


Using Gradient 

```{r}

# gradient
  Call.g2 <- ggplot(Call911.short2[year==2017]) +
                stat_density_2d(aes(x=Longitude, y=Latitude, fill = ..level..,  alpha =..level..), 
                    size = 1, bins = 16, geom = "polygon") +
                scale_fill_gradient(low = "green", high = "red") +
                geom_path(data=Seattle.points, aes(long, lat, group = group), colour = "#d3d3d3") +
                ggthemes::theme_map() +
                theme(plot.background = element_rect(fill = "black"),
                      legend.background = element_rect(fill = "black"),
                      legend.position = "right",
                      legend.key = element_rect(fill = "black"),
                      legend.text = element_text(color = "white")) +
                coord_map() +
                xlim(Seattle@bbox[1,1], Seattle@bbox[1,2]) +
                ylim(Seattle@bbox[2,1], Seattle@bbox[2,2])
  
  Call.g2
  ggplotly(Call.g2) %>% layout(plot_bgcolor="black")
 
```


Let's try to compare density for each year


```{r}

# gradient
  Call.g3 <- ggplot(Call911.short2) +
                stat_density_2d(aes(x=Longitude, y=Latitude, frame=year, fill = ..level..,  alpha =..level..), 
                size = 1, bins = 16, geom = "polygon") +
                scale_fill_gradient(low = "green", high = "red") +
                geom_path(data=Seattle.points, aes(long, lat, group = group), colour = "#d3d3d3") +
                ggthemes::theme_map() +
                theme(plot.background = element_rect(fill = "black"),
                      legend.background = element_rect(fill = "black"),
                      legend.position = "right",
                      legend.key = element_rect(fill = "black"),
                      legend.text = element_text(color = "white"),
                      title = element_text(color = "white"))  +
                coord_map() +
                xlim(-122.36, -122.30) +
                ylim(47.58, 47.68) 
    
  ggplotly(Call.g3) %>% layout(plot_bgcolor="black", xaxis = list(range = c(-122.36, -122.30)),
                               yaxis = list(range(c(47.58, 47.68))))
  
 
```


Animation per month

```{r, fig.show="animate"}

p <- ggplot() +
      geom_path(data=Seattle.points, aes(long, lat, group = group), colour = "#d3d3d3") +
      geom_point(data = Call911.short2[year==2016], aes(x=Longitude, y=Latitude, color = Type, frame = month)) +
      ggthemes::theme_map() +
              theme(plot.background = element_rect(fill = "black"),
                    legend.background = element_rect(fill = "black"),
                    legend.position = "right",
                    legend.key = element_rect(fill = "black"),
                    legend.text = element_text(color = "white", size=12),
                    title = element_text(color = "white", size =12))  +
              coord_map() +
              xlim(-122.5, -122.2) +
              ylim(47.5, 47.7) 

gganimate(p)
```


Just to check, you can fix scales in ggplot

```{r}
  for (i in (2010:2017)) {
  
    plot <- ggplot(Call911.short2[year==i]) +
            stat_density_2d(aes(x=Longitude, y=Latitude, fill = ..level..,  alpha =..level..), 
                            size = 1, bins = 16, geom = "polygon") +
            scale_fill_gradientn(colours = c("green", "yellow","red"),
                                 limits=c(0, 5000) ) +
            geom_path(data=Seattle.points, aes(long, lat, group = group), colour = "#d3d3d3") +
            ggthemes::theme_map() +
            theme(plot.background = element_rect(fill = "black"),
                  legend.background = element_rect(fill = "black"),
                  legend.position = "right",
                  legend.key = element_rect(fill = "black"),
                  legend.text = element_text(color = "white"),
                  title = element_text(color = "white")) +
            coord_map() +
            xlim(-122.36, -122.30) +
            ylim(47.58, 47.68)  + 
            ggtitle(paste("Year", i))
    print(plot)
  }
```


