---
title: "Data Visualization with R - Exercice 1"
author: "AGPC - Flec"
date: "Novembre 2017"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    theme: united
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(data.table)
library(scales)
library(ggpubr)
```


```{r, echo=FALSE}

## data
  reportDF <- read.csv2(file = "./datasets/report2017.csv")
  reportDF$date <- as.Date(reportDF$date)
  reportDF$portfolio <- as.numeric(reportDF$portfolio)


  DFagg <- as.data.table(
          reportDF %>%
           group_by(LOB, network, date) %>%
           summarize(NB = sum(newBusiness),
                     canc = sum(cancellation),
                     portfolio = sum(portfolio)) %>%
           mutate(NB.rate = NB/portfolio,
                  canc.rate = canc/portfolio,
                  netInflow = as.numeric(NB - canc)) %>%
           select(-c(NB, canc))
  )
 
```

### EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX  
### EXERCICE 1  
### EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX  

Using ggplot, let's plot for **Agent** network
- per LOB (Motor = solid, Household = dashed)
- per date in x-axis
3 KPIs : 
- New Business Rate : green line
- Cancellation Rate : red line
- NetInflow : bar plot

**Caution** : remember that ggplot does not allow to plot chart with different y-axis ni the same area !


### Solution : 

1/ Cancellation and NB rate using geom_line : 

Generic ggplot theme

```{r}
## generic theme  :
  myTheme <-  function(x) { theme(axis.text.x = element_text(angle = -90),                  
                                  axis.title.x = element_blank(),
                                  axis.title.y = element_blank(),
                                  legend.position = "top")
  }
```

### SOLUTION 1 : Best in class Solution : using tidy dataframe to smooth ggplot use


1/ Cancellation and NB rate using geom_line : 

1. Create Dataset
```{r}
  DF.tidy <- DFagg %>% 
             tidyr::gather(KPI, value, -date, -LOB, -network) %>% 
             filter(network == "Agent") %>% 
             as.data.table()

```


2. Call ggplot on the tidy dataframe
```{r}

 a1 <-   ggplot(DF.tidy[KPI %in% c("NB.rate", "canc.rate")]) +                        # data
         geom_line(aes(x=date, y=value, color = KPI)) +                               # geoms
         facet_wrap(~LOB, dir = "h") +                                                # facet for LOB
         scale_y_continuous(labels = scales::percent ) +                              # Scale Y
         scale_x_date(date_labels = "%b %y", date_breaks = "3 months") +              # Scale X
         scale_colour_manual(name="KPI's", 
                             values=c("NB.rate" = "green", "canc.rate" = "red"),
                             label = c("NB.rate"="New Business", "canc.rate"="Cancellation")) +
        myTheme() + ggtitle("Evolution of New Business/Cancellation rate per Line of Business")   # Theme
 a1
 
```
 
 
### SOLUTION 2 : Less adviced : using dataframe as it is

1. same data
2. ggplot on the untidy dataframe
```{r}
 b1 <-  ggplot(DFagg[network == "Agent"]) +                                           # data
        geom_line(aes(x=date, y=NB.rate, color = "NB")) +                             # geom NB
        geom_line(aes(x=date, y=canc.rate, color = "Canc")) +                         # geom Cancellation
        facet_wrap(~LOB, dir="h") +                                                   # facet for LOB
        scale_y_continuous(labels = scales::percent ) +                               # Scale Y
        scale_x_date(date_labels = "%b %y", date_breaks = "3 months") +               # Scale X
        scale_color_manual(name="KPI's", 
                            values= c(NB = "#00FF00", Canc = "#FF0000"),
                            label = c(NB = "New Business", Canc = "Cancellation")) +
        myTheme() + ggtitle("Evolution of New Business/Cancellation rate per Line of Business")   # Theme
b1    

```


### SOLUTION 3 : Solution 2 wrapped in a function 

Note that to map ggplot in a function, some adjustments need to be done : 
- aes => aes_string if the function variables appear in aes()
- Quoted string need to be wrapped with shQuote()
- as.formula for facet options

```{r}

   plot.NB.canc <- function(DF, x.date, y.nb, y.canc, grp) {
        ggplot(DF) + 
          geom_line(aes_string(x=x.date, y=y.nb, color = shQuote("NB"))) +
          geom_line(aes_string(x=x.date, y=y.canc, color = shQuote("Canc"))) +
          facet_wrap(as.formula(paste0("~",grp)), dir="h") +
          scale_y_continuous(labels = scales::percent ) +
          scale_x_date(date_labels = "%b %y", date_breaks = "3 months") +
          scale_color_manual(name="KPI's", 
                              values= c(NB = "#00FF00", Canc = "#FF0000"),
                              label = c(NB = "New Business", Canc = "Cancellation")) +
          myTheme() + ggtitle("Evolution of New Business/Cancellation rate per Line of Business")   
      }    
   c1 <- plot.NB.canc(DFagg[network == "Agent"], "date", "NB.rate", "canc.rate", "LOB")
   c1
```


2/ Adding NetInflow bar chart and combine all.

=> Here follows the global solution (I choose the tidy method embedded in a function)


```{r}
## generic theme  :
  myTheme <-  function(x) { theme(axis.text.x = element_text(angle = 90),                  
                                  axis.title.x = element_blank(),
                                  axis.title.y = element_blank(),
                                  legend.position = "right")
  }

## tidy dataframe
  DF.tidy <- DFagg %>% tidyr::gather(KPI, value, -date, -LOB, -network) %>% filter(network == "Agent") %>% as.data.table()


## function 1 : line 
 g1.f <- function(df, grp)  {ggplot(df) +                                                           # data
                       geom_line(aes(x=date, y=value, color = KPI)) +                               # geoms
                       facet_wrap(as.formula(paste0("~",grp)), dir="h") +                           # facet for LOB
                       scale_y_continuous(labels = scales::percent ) +                              # Scale Y
                       scale_x_date(date_labels = "%b %y", date_breaks = "3 months") +              # Scale X
                       scale_colour_manual(name="KPI",                                              # Scale Color
                                           values=c("NB.rate" = "green", "canc.rate" = "red"),
                                           label = c("NB.rate"="New Business", "canc.rate"="Cancellation")) +
                      myTheme()                                                                     # Theme
 }
 ## function 2 : bar  
 g2.f <- function(df, grp)  {ggplot(df) +                                                           # data
                       geom_bar(aes(x=date, y=value, fill = KPI),                                   # geoms
                                stat = "identity", position = "dodge") +                            # 
                       facet_wrap(as.formula(paste0("~",grp)), dir="h") +                           # facet for LOB
                       scale_y_continuous(labels = scales::comma ) +                                # Scale Y
                       scale_x_date(date_labels = "%b %y", date_breaks = "3 months") +              # Scale X
                       scale_fill_manual(name = "KPI",
                                         values=c("netInflow" = "khaki"),                           # Scale Fill
                                         label = c("netInflow"="Net Inflow")) +
                      myTheme()                                                                     # Theme
 }
 
 
 g1 <- g1.f(DF.tidy[KPI %in% c("NB.rate", "canc.rate")], "LOB")
 g2 <- g2.f(DF.tidy[KPI == "netInflow"], "LOB") 

ggpubr::ggarrange(g1, g2, align = "v", nrow = 2)


```

