---
title: "Data Visualization with R - Exercice 3"
author: "AGPC - Flec"
date: "Novembre 2017"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    theme: united
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(data.table)
library(scales)
library(plotly)
```


```{r, echo=FALSE}

## data
  reportDF <- read.csv2(file = "./datasets/report2017.csv")
  reportDF$date <- as.Date(reportDF$date)
  reportDF$portfolio <- as.numeric(reportDF$portfolio)


  DFagg <- as.data.table(
          reportDF %>%
           group_by(LOB, network, date) %>%
           summarize(NB = sum(newBusiness),
                     canc = sum(cancellation),
                     portfolio = sum(portfolio)) %>%
           mutate(NB.rate = NB/portfolio,
                  canc.rate = canc/portfolio,
                  netInflow = as.numeric(NB - canc)) %>%
           select(-c(NB, canc))
  )
 
```

### EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX  
### EXERCICE 3 
### EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX EX  

Try to reproduce at least one of New Business, Cancellation and Net Inflow charts with plot_ly

Nota : with plotly, you can overlay all sort of chart, even on several y axis

```{r}

plot1 <- plot_ly(DFagg[network == "Agent"]) %>%
            add_trace( x = ~date, y = ~-netInflow, 
                       type = 'bar', 
                       yaxis = "y2",
                       opacity = .4, 
                       color = ~LOB, 
                       colors = colorRamp(c("steelblue", "grey")),
                       hoverinfo = 'text', 
                       text = ~paste('Net Inflow : ', comma(netInflow),'<br> LOB: ', LOB)) %>% 
            add_trace( x = ~date, y = ~NB.rate*100, 
                       name = '% New Business', 
                       type = 'scatter',  
                       mode = 'lines', 
                       linetype = ~LOB, 
                       line = list(color = "green"), 
                       hoverinfo = 'text', 
                       text = ~paste('% New Business : ', round(NB.rate*100, 1), "% ",'<br> LOB: ', LOB))  %>% 
            add_trace( x = ~date, y = ~canc.rate*100, 
                       name = '% Cancellation',  
                       type = 'scatter', 
                       mode = 'lines+markers', 
                       linetype = ~LOB,
                       line = list(color = "red"),
                       hoverinfo = 'text', 
                       text = ~paste('% Cancellation : ', round(canc.rate*100, 1), "% ", '<br> LOB: ', LOB)) %>% 
            layout( title = "Evolution of New Business and Cancellation rates per Line of Business",
                       barmode = 'dodge',
                       yaxis = list(ticksuffix = '%', 
                                    title = "(in % of portfolio N)",
                                    tickfont = list(size = 10)),
                       yaxis2 = list(overlaying = "y",
                                    side = "right",
                                    tickfont = list(size = 10),
                                    showgrid = FALSE),
                       xaxis = list(tickangle = -45, 
                                    title = "",
                                    tickfont = list(size = 10)),
                       margin = list(b=70),
                       legend = list( x = 1.05,                        # avoid overlap of legend and y-axis
                                      y = 0.5, yanchor = "center") )   # valign

plot1 
```


Technically feasible but not very lisible : it is adviced to create several plots.


```{r}
# choose your colors per LOB for instance
  cols <- data.table(v = levels(DFagg$LOB), colors = c("steelblue", "#dee5e4"))
  
  plotly.nbcc <- function(df, ntw, lob, legend) {
    
        # NB and Cancellation
        p.nbcc <- plot_ly(df[network == ntw & LOB == lob]) %>%
            add_lines(x = ~date, y = ~canc.rate*100, 
                      name = '% Cancellation',  
                      line = list(color = "red", dash = "dot"),
                      showlegend = legend,
                      hoverinfo = 'text', 
                      text = ~paste('% Cancellation : ', round(canc.rate*100, 1), "% ", '<br> (', ntw, '-', lob, ')')) %>%
            add_lines(x = ~date, y = ~NB.rate*100, 
                      name = '% NewBusiness',  
                      line = list(color = "green"),
                      showlegend = legend,
                      hoverinfo = 'text', 
                      text = ~paste('% New Business : ', round(NB.rate*100, 1), "% ", '<br> (', ntw, '-', lob, ')')) %>%
            layout(xaxis = list(title = ""),
                   yaxis = list(ticksuffix = '%', 
                                tickfont = list(size = 8)))
  }
  
  plotly.ni <- function(df, ntw, lob, legend) {
      # Net Inflow
      p.ni <- plot_ly(df[network == ntw & LOB == lob],
                x = ~date, y = ~netInflow, 
                type = 'bar', 
                name = "Net Inflow",
                marker = list(color = cols[v==lob, colors] ) ,
                showlegend = legend,
                opacity = .6,
                hoverinfo = 'text', 
                text = ~paste('Net Inflow : ', comma(netInflow), '<br> (', ntw, '-', lob, ')')) %>%
          layout(xaxis = list(tickangle = -90,
                              side = "top",
                              title = "",
                              autotick = F, 
                              dtick = "M2",
                              tickfont = list(size = 8)),
                 yaxis = list(tickfont = list(size = 8)))

 }
  
  
  
```

And see the results 

```{r}


 ## paste plots together
 plotly.KPI <- function(df, ntw, lob, legend) {
 
    subplot(plotly.nbcc(df, ntw, lob, legend), 
            plotly.ni(df, ntw, lob, legend), 
            nrows = 2, shareX = TRUE, margin = .08) %>%
    layout( title = paste("Net Inflow, New Business & Cancellation rates  : <b>", ntw, "-", lob, "</b>"),
            margin = list(b=70),
            legend = list( x = 1.05,                        # avoid overlap of legend and y-axis
                           y = 0.5, yanchor = "center") )   # valign
  
 }

 plotly.KPI(DFagg, "Agent", "Motor", T)
 
```


Use subplot to gather charts :

```{r}

   subplot(plotly.nbcc(DFagg, "Agent", "Motor", T), 
           plotly.nbcc(DFagg, "Direct", "Motor", F),
           plotly.ni(DFagg, "Agent", "Motor", T), 
           plotly.ni(DFagg, "Direct", "Motor", F), 
            nrows = 2, shareX = TRUE,shareY = T, margin = .08) %>%
    layout( title = paste("Net Inflow, New Business & Cancellation rates  : <b> Agent vs Direct - Motor</b>"),
            legend = list( x = 1.05,                        # avoid overlap of legend and y-axis
                           y = 0.5, yanchor = "center") )   # valign
  
```

